Staged Abstract Interpreters

### 2018-11-03 paper/10_intro.tex
fix typos in paper

Multi-stage programming has been widely used to improve the performance in many domains, 
such as optimizing compilers and domain-specific languages \cite{DBLP:conf/pldi/RompfSBLCO14, DBLP:conf/snapl/RompfBLSJAOSKDK15,
DBLP:journals/tecs/SujeethBLRCOO14, DBLP:conf/gpce/SujeethGBLROO13, DBLP:journals/jfp/CaretteKS09},
numerical computation \cite{PGL-038, DBLP:conf/pepm/AktemurKKS13}, 
generic programming \cite{DBLP:journals/pacmpl/Yallop17, Ofenbeck:2017:SGP:3136040.3136060}, 
data processing \cite{DBLP:conf/oopsla/JonnalageddaCSRO14, DBLP:conf/popl/KiselyovBPS17}, 
query compilation in databases \cite{DBLP:conf/osdi/EssertelTDBOR18, DBLP:conf/sigmod/TahboubER18},
etc.
Likewise, specializing static analyses by partial evaluation had been emerged in late 90s 
\cite{damian1999partial, amtoft1999partial, Boucher:1996:ACN:647473.727587, ashley:practical}, 
and indeed it is able to effectively remove the interpretive overhead of 
repeatedly traversing the abstract syntax tree. However, it is still worth 
to investigate the idea deeper from a modern perspective based on meta-programming 
for general abstract interpreters. \note{TR: sounds weak, what is new?}
Practically, precise whole-program analyses formulated as abstract interpretation 
can be very slow due to the inherent complexity of the algorithm and the large scale 
of the analyzed programs \cite{toman_et_al:LIPIcs:2017:7121} \todo{other cite?};
particularly, we would like a systematic approach to optimize program 
analyzers and meanwhile minimally modifying the analyzer programs.
In the present work, we show how to achieve this goal by bring existing 
high-performance computing facilities, in the style of multi-stage generative
programming, to abstract interpretation semantic artifacts, i.e., 
abstract interpreters.

The slogan of multi-stage programming is "abstraction without regret". 
Of course, all interpreters are metalinguistic abstractions, but some interpreters are more "abstract"
than others \todo{maybe rephrase}. 
One technical challenge is to deal with the non-determinisism, fixed-point iterations and abstract 
domains in abstract interpreters when staging is introduced. The abstract interpreter, as a semantic artifact, 
should be written in a style that is easy for people to communicate the formulation, but also be friendly
to specialization and generating efficient code. \todo{We discuss the form of abstract interpreters for staging}
We show the use LMS framework, that enables multi-stage programming by just changing the types 
of the analyzer program, and we thus bridge the gap between designing sound program analyses 
and implementing efficient program analyzers. \note{TR: a lot of back and forth here, where is story going?}

>>>>

Multi-stage programming has been widely used to improve the performance in many domains, 
such as optimizing compilers and domain-specific languages \cite{DBLP:conf/pldi/RompfSBLCO14, DBLP:conf/snapl/RompfBLSJAOSKDK15,
DBLP:journals/tecs/SujeethBLRCOO14, DBLP:conf/gpce/SujeethGBLROO13, DBLP:journals/jfp/CaretteKS09},
numerical computation \cite{PGL-038, DBLP:conf/pepm/AktemurKKS13}, 
generic programming \cite{DBLP:journals/pacmpl/Yallop17, Ofenbeck:2017:SGP:3136040.3136060}, 
data processing \cite{DBLP:conf/oopsla/JonnalageddaCSRO14, DBLP:conf/popl/KiselyovBPS17}, 
query compilation in databases \cite{DBLP:conf/osdi/EssertelTDBOR18, DBLP:conf/sigmod/TahboubER18},
etc.
Likewise, specializing static analyses by partial evaluation had been emerged in late 90s 
\cite{damian1999partial, amtoft1999partial, Boucher:1996:ACN:647473.727587, ashley:practical}, 
and indeed it is able to effectively remove the interpretive overhead of 
repeatedly traversing the abstract syntax tree. However, it is still worth 
to investigate the idea deeper from a modern perspective based on meta-programming 
for general abstract interpreters. \note{TR: sounds weak, what is new?}
Practically, precise whole-program analyses formulated as abstract interpretation 
can be very slow due to the inherent complexity of the algorithm and the large scale 
of the analyzed programs \cite{toman_et_al:LIPIcs:2017:7121} \todo{other cite?};
particularly, we would like a systematic approach to optimize program 
analyzers and meanwhile minimally modifying the analyzer programs.
In the present work, we show how to achieve this goal by bringing existing 
high-performance computing facilities, in the style of multi-stage generative
programming, to abstract interpretation semantic artifacts, i.e., 
abstract interpreters.

The slogan of multi-stage programming is "abstraction without regret". 
Of course, all interpreters are metalinguistic abstractions, but some interpreters are more "abstract"
than others \todo{maybe rephrase}. 
One technical challenge is to deal with the non-determinisism, fixed-point iterations and abstract 
domains in abstract interpreters when staging is introduced. The abstract interpreter, as a semantic artifact, 
should be written in a style that is easy for people to communicate the formulation, but also be friendly
to specialization and generating efficient code. \todo{We discuss the form of abstract interpreters for staging}
We show the usage of the LMS framework, that enables multi-stage programming by just changing the types 
of the analyzer program, and we thus bridge the gap between designing sound program analyses 
and implementing efficient program analyzers. \note{TR: a lot of back and forth here, where is story going?}
